---
title: "linreg algorithm - R implementation"
author: "Michal Burdukiewicz"
date: "May 29, 2015"
output:
  html_document:
    toc: yes
---


# Questions

## Diagram questions

Naming convention: Q, number of part (roman numerals), underscore, number of question on diagram (greek numerals).

```{r,echo=FALSE,message=FALSE,results='asis'}
#Each question must start with Q/[no space/]numberOfPart_numberOfQuestion.
#Only one question per line.
source("extract_questions.R")

#cat(extract_questions("linreg_scheme.Rmd"))

```

## Answered

Q0_1. Linreg manual indicates that apply baseline means substract baseline. If the baseline is smaller than 0, should it be also substracted?  
Q0-1: apply baseline is indeed always subtract baseline. So when the baseline becomes negative you are in fact adding to the observed fluorescence values. 

Q0_2. If the minimum value is 0, should I add a small epsilon to fluorescence values?  
Q0-2: I think that you mean observed minimum fluorescence is 0. In that case the qPCR machine has done a baseline subtraction. When that is a constant baseline you can add any constant you want to keep all data in the analysis. When the machine does a baseline trend you have to go back to the machine and get the raw data.

Q0_3. Which algorithm was used for numerical derivation?  
Q0-3: I simply took the difference of two values or the difference of two of those difference. The purpose of the procedure is not to determine the exact SDM value, but only to identify the cycle that marks the start of the plateau.

QI_1. How to define start of the exponential phase if there are no detectable jump?  
Q1-1: the exponential phase starts after the ground phase noise. So there must always be a jump. I have seen dat from a qPCR machine that were not raw, but fitted to some kind of curve. In that case the noise is lost and thus the jump is lost and LinRegPCR fails.

QI_2. How to set a baseline as the average of the 6th and the 7th point below the plateau phase when plateau phase is not visible?  
Q1-2: When the plateau phase, as the phase after the SDM cycle, is not there, LinRegPCR flags the sample as 'no plateau' and does give a N0 value only when the PCR efficiency is not deviating from the other samples in which the same amplicon is amplified.

QII_1. What if step is negative? should it be added or substracted?  
Q2-1: See Q0-1: the step value is always subtracted: so when it becomes negative you are in fact adding something.

QII_2. What means 2.step?  
Q2-2: 2.step means 2 x step or in words: two times the step value. It is in the procedure to go 2 steps back when the shape of the curve chances fron convex to concave (or v.v.). This is to make sure that the first step/2 that is subtracted again does leave the procedure in the same 'shape'. 



QADD1: From manual: "Corrected a bug that occurred when the plateau phase data were not fully continuously increasing. This then resulted in a two plateau phase points being selected and the sample being excluded. Note that this correction can lead to changed results when data are re-analysed." - how bug was corrected?

A: "The correction in the code of LinRegPCR you refer to in the Additional Question was made in September 2011 and was required because of something that occurred in the datasets of the Methods paper (competimer set, samples G_D_1 and F_C_2). In those sets the maximum of the first derivative (FDM) was followed by a dip in the fluorescence values.  
Why was that important in the code: the second derivative maximum (SDM) is always earlier than the FDM. So to save processing time, the search for the SDM was limited from the ‘jump’ as the last noisy cycle to the FDM+2. The plus 2 is to allow the first and second difference in fluorescence values to be calculated. However, in case of a ‘dip’ after FDM, this resulted an exponential phase of 2 points. So, the solution was to remove the +2 in case there was no continuous increase of 2 cycles after the FDM (less increase than before FDM but still increasing)."


# List of abbrevations

* BL: baseline.
* S_L: slope of points in the lower half of the exponential phase.
* S_U: slope of points in the upper half of the exponential phase.
* SDM: second derivative maximum, end of the exponential phase.
* Start of the exponential phase: a cycle C for which (counting from plateau) fluorescence value is bigger than in cycle C + 1.


# Baselining


## Part 0

```{r,echo=FALSE}
library(DiagrammeR)
diagram <- "
graph TB
subgraph linreg
A[raw sample]==>B{set baseline to <br/> minimum observation}
B== BL != 0 ==>C[substract baseline]
C==>D{check for amplification <br/> Samples are skipped when less than seven times <br/> increase in fluorescence values is observed}
D== yes ==>Eyes[Determine SDM cycle: <br/> end of the exponential phase <br/> the approximate derivative is a difference of two values]
D== no ==>Eno[skip sample]
Eyes==>F[Move to Part I]
B== BL != 0 ==>Cbis[if the observed minimum fluorescence is 0 <br/> and machine BL is constant, add any value. <br/> Get raw values in case of baseline trend]

style A fill:#DCEBE3
style B fill:#77DFC9
style C fill:#DEDBBA
style Cbis fill:#DEDBBA
style D fill:#77DFC9
style Eyes fill:#DEDBBA
style Eno fill:#DEDBBA
end

subgraph questions
A2[no questions]==>B2[no questions]
B2==>C2[no questions]
C2==>D2[no questions]
D2== yes ==>Eyes2[no questions]

D2== no ==>Eno2[no questions]

style A2 fill:#DCEBE3
style B2 fill:#77DFC9
style C2 fill:#DEDBBA
style D2 fill:#77DFC9
style Eyes2 fill:#DEDBBA
style Eno2 fill:#DEDBBA
end
"

mermaid(diagram)
```

## Part I

```{r,echo=FALSE}

diagram <- "
graph TB
subgraph linreg
START[Move from part 0]==>A[Determine start of the exponential phase]
A==>B[set baseline too high: <br/> baseline is the average of the 6th and the 7th point <br/> below the plateau phase]
B==>C[apply baseline]
C==>D{compare S_U and S_L <br/> When the exponential phase has an uneven number of points, <br/> the middle point is in the top as well as the bottom part}
F== no ==>C
D== S_U > S_L ==>Esmaller[Define step: 0.005*baseline]
D== S_U < S_L ==>Ebigger[Decrease baseline by 0.01]
Ebigger==>F{baseline < min. observ}
F== yes ==>G[baseline error]
Esmaller==>H[Move to part II]

style A fill:#77DFC9
style B fill:#DEDBBA
style C fill:#77DFC9
style D fill:#DEDBBA
style Ebigger fill:#77DFC9
style Esmaller fill:#77DFC9
style F fill:#DEDBBA
style G fill:#77DFC9
end

subgraph questions
START2[Move from part 0]==>A2[if no jump, linreg fails]
A2==>B2[cannot set baseline without plateau]
B2==>C2[no questions]
C2==>D2[no questions]
D2== S_U > S_L ==>Esmaller2[no questions]
D2== S_U < S_L ==>Ebigger2[no questions]
Ebigger2==>F2[no questions]
F2== no ==>C2
F2== yes ==>G2[no questions]

style A2 fill:#77DFC9
style B2 fill:#DEDBBA
style C2 fill:#77DFC9
style D2 fill:#DEDBBA
style Ebigger2 fill:#77DFC9
style Esmaller2 fill:#77DFC9
style F2 fill:#DEDBBA
style G2 fill:#77DFC9
end
"

mermaid(diagram)
```


## Part II

```{r,echo=FALSE}

diagram <- "
graph TB
subgraph linreg
START[Move from part I]==>A[BL = BL + step]
A==>B[apply baseline]
B==>C{compare S_U and S_L}
Dbigger[BL = BL - 2.step <br/> step = 0.5*step]==>B
C==S_U - S_L < 1e-05==>E[Baseline corrected sample]
C== S_U < S_L ==>Dbigger
C== S_U > S_L ==>A

style A fill:#77DFC9
style B fill:#DEDBBA
style C fill:#77DFC9
style Dbigger fill:#DEDBBA
style E fill:#DEDBBA
end

subgraph questions
START2[Move from part I]==>A2[Always substract step]
A2==>B2[no questions]
B2==>C2[no questions]
C2== S_U < S_L ==>Dbigger2[step times 2]
Dbigger2==>B2
C2== S_U - S_L < 1e-05 ==>E2[no questions]
C2== S_U > S_L ==>A2

style A2 fill:#77DFC9
  style B2 fill:#DEDBBA
  style C2 fill:#77DFC9
  style Dbigger2 fill:#DEDBBA
  style E2 fill:#DEDBBA
  end
"

mermaid(diagram)
```

